---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document: default
  pdf_document: default
runtime: knitr
keep_md: yes
---
## Author : Nima Arvin

## setting the global options to show the code and results for all code chunks
```{r}
knitr::opts_chunk$set(echo = TRUE)
```







## Code for reading in the dataset and/or processing the data

```{r Loading the data , cache=TRUE}
library(knitr)
library(lubridate)
wd <- getwd()
zipfile <-download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",destfile=paste0(wd,"/file.zip")) 
unzip(paste0(wd,"/file.zip"))
maindata <- read.csv(paste0(wd,"/activity.csv"), header = TRUE)
maindata$date <- ymd(maindata$date)
summary(maindata)
str(maindata)
```


## Daily Activity Pattern

### What is mean total number of steps taken per day?
```{r}
library(dplyr)
gdata <- group_by(maindata, date)
dailymeans<- summarise(gdata, mean=mean(steps, na.rm = TRUE))
dailymeans
```

### Median number of steps taken each day
```{r}
dailymedians<- summarise(gdata, mean=median(steps, na.rm = TRUE))
dailymedians
```

### Total number of steps taken per day
```{r}
dailytotals <- summarise(gdata, totals=sum(steps, na.rm = TRUE))
dailytotals
```

### Histogram of the total number of steps taken each day
```{r}
with(dailytotals,hist(totals, col = "green"))
```

### Mean and Median of the total number of steps taken per day
```{r}
### Mean of total number of steps per day
mean(dailytotals$totals)

### Median of total number of steps per day
median(dailytotals$totals)

```



### Time series plot of the 5 minutes interval

```{r}
library(dplyr)
library(ggplot2)
gidata <- group_by(maindata, interval)
intervalmeans <- summarize(gidata, intmeans=mean(steps, na.rm = TRUE))
intervalmeans

g2 <- ggplot(intervalmeans, aes(interval, intmeans, group=1))
g2+geom_point(col="black", size=1)+geom_line(col="red")+labs(title="Mean of number of steps per intervals", x="Intervals observed", y="Mean of number of steps")


```

### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
intervalmeans[intervalmeans[,2]==max(intervalmeans[,2]),]
```



### Time series plot of the average number of steps taken
```{r}
g <- ggplot(dailymeans, aes(date,mean, group=1))
g + geom_point(col="red")+ geom_line()+geom_smooth(method = "glm", na.rm = TRUE)+ theme(axis.text.x = element_text(angle = 90, hjust = 1))+labs(title="Average Steps Per Day",x="Observation Date",y="Mean of number of steps" )
```

### Are there differences in activity patterns between weekdays and weekends?
```{r}
library(lubridate)
dailytotals$weekday <- wday(dailytotals$date, label = TRUE)
wdg <- group_by(dailytotals, weekday)
weekdayavg <- summarize(wdg, WeekdayAverage=mean(totals, na.rm = TRUE))
g3 <- ggplot(weekdayavg, aes(weekday,WeekdayAverage, group=1))
g3+geom_point(col="black", size=2)+geom_line(col="red")
```


### How did the weekdays trend looked like over time ?
```{r}
dailytotals$weekday <- wday(dailytotals$date, label = TRUE)
g1 <- ggplot(dailytotals, aes(date, totals, group=weekday))
g1+geom_point(size=1, col="black")+geom_line(col="red")+facet_grid(.~weekday)+theme(axis.text.x = element_text(angle = 90, hjust = 1))+geom_smooth(method = "lm")
```


### Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
day <- weekdays(maindata$date)
daylevel <- vector()
for (i in 1:nrow(maindata)) {
    if (day[i] == "Saturday") {
        daylevel[i] <- "Weekend"
    } else if (day[i] == "Sunday") {
        daylevel[i] <- "Weekend"
    } else {
        daylevel[i] <- "Weekday"
    }
}
maindata$daylevel <- daylevel
maindata$daylevel <- factor(maindata$daylevel)

gint <- group_by(maindata, daylevel, interval)
aint <- summarize(gint,ave=mean(steps, na.rm = TRUE) )


g4 <- ggplot(aint, aes(interval, ave, group=1))
g4+geom_point(size=1, col="red")+ geom_line(col="black")+facet_grid(daylevel~.)



```



## Imputing missing values

### Calculate and report the total number of lines with missing values in the dataset
```{r}
sum(is.na(maindata))
```

### Get a report of the columns that have missing values
```{r}
summary(is.na(maindata))
```



### Code to describe and show a strategy for imputing missing data

1. The strategy will be to substitude the missing value with the mean of the number of steps for that interval
```{r}
z <- aggregate(steps~ interval, data = maindata,mean)
for (i in 1:nrow(maindata)) {
    if (is.na(maindata$steps[i])) {
        maindata$steps[i] <- z[z[,1]==maindata$interval[i],2]
    }
}
maindata.no.missing <- maindata
summary(is.na(maindata.no.missing))

```


### Histogram of the total number of steps taken each day after missing values are imputed
```{r}

ggroup <- summarise(group_by(maindata.no.missing, date),sum=sum(steps, na.rm = TRUE))
with(ggroup, hist(sum, col = "red"))

```



